MetaTracts - A Method for Robust Extraction and Visualization of Carbon Fiber Bundles in Fiber Reinforced Composites

### 2014-09-22 chapters/MetaTracts.tex

\begin{figure}[tb]
\centering
	%\includegraphics[width=0.45\textwidth]{imagesMT2014/crop-16/length_dist.png}
		\includegraphics[width=0.45\textwidth]{imagesMT2014/image5}
	\caption{(A) MetaTracts colored according to mean orientations. (B) length distribution of individual MetaTracts for a particular bundle.}
	\label{fig:length_distribution}
\end{figure}

@@

The directed Hausdorff distance is very sensitive to outliers in the data.
However, because MetaTracts after orientation clustering are constructed using cylinders with similar orientations, they are not plagued by outliers.
To cluster based on MetaTract proximity, we used single linkage hierarchical clustering.
Hierarchical clustering has a single parameter $h$, the desired number of clusters.
In hierarchical clustering each objects starts in its own cluster and clusters are merged based on some criterion.
Clusters are merged until there are only $h$ clusters left.
Hierarchical clustering is intuitive since it is easy to trace how clusters are formed and merged.
Single linkage clustering finds pairs of objects $p \in P$ and $q \in Q$ where $P \neq Q$ which are closer than other such pairs, and merges the containing clusters $P$ and $Q$.
We found that single linkage hierarchical clustering had two major drawbacks.
First, the clustering would produce some small clusters of just a few MetaTracts.
These MetaTracts were anomalies caused by noise in the data and did not represent true fiber bundles.
Second, if two fiber bundles were parallel for some of their length and then separated, they would sometimes be clustered into the same fiber bundle.
A short MetaTract which was parallel to both and did not extend into the separation region could form a link between the two fiber bundles, causing them to be clustered into a single bundle.
To address the problem of small clusters, we applied hierarchical clustering and then identified small clusters with few MetaTracts.
We removed the MetaTracts that were in those clusters from the data set and reapplied hierarchical clustering.
To address the problem of short MetaTracts joining different fiber bundles, we applied hierarchical clustering and then removed the
shortest fibers (fibers less than 0.6 times the median length) in each bundle. We then reapplied hierarchical clustering.
We repeated both until a steady state of clusters is reached and no new small fibers can be removed. The results of clustering the orientation clusters Fig.~\ref{fig:orientation_clustering}(A,C) are shown in Fig.~\ref{fig:orientation_clustering}(D,E) respectively. Note how the purple cluster Fig.~\ref{fig:orientation_clustering}(E) bundle ``7'' is extracted well.
\begin{figure}
\centering
		%\includegraphics[width=0.4\textwidth]{imagesMT2014/garb1}~
		\includegraphics[width=0.5\textwidth, trim = 0mm 70mm 0mm 15mm, clip]{imagesMT2014/image7b.pdf}
	%\includegraphics[width=0.45\textwidth]{imagesMT2014/crop-16/bundles}~ 
	%\includegraphics[height=0.35\textwidth]{imagesMT2014/crop-16/vol_mesh.png} 
	\caption{(A) Volume rendering of data set Figure~\ref{fig:data-char}A.(B) shows a single slice of the data set. (C) shows two of the extracted meshes together. (D) shows the meshes separately.}
	\label{fig:crop-16-decomp}
\end{figure}

@@

\section{Parameter choices}
\label{sec:param_choices}
\begin{figure}
\centering
	\includegraphics[width=0.4\textwidth]{imagesMT2014/crop-16/histogram_10_A.PNG}
	\caption{Number of tracts and median, minimum and maximum length of individual MetaTracts in the orientation cluster Fig.~\ref{fig:orientation_clustering}, clustered into 10 clusters (Fig.~\ref{fig:crop-16-decomp}A top). The unit for length is the grid cube length. Figure~\ref{fig:length_distribution}(B) shows the length of individual MetaTracts in cluster 6. Clusters labeled 2,3,4,5 and 7 all have cardinality less than 10. }
	\label{fig:len_dist_crop16}
\end{figure}
The critical parameters are $K$ for the K-means in orientation based clustering and $h$ for the hierarchical clustering. K denotes the number of major directions, which is either known a priori or can be easily estimated by looking at the weaving pattern. 
Our framework is robust to the choice of $h$ (Section~\ref{sec:results}). Large $h$ keeps the major bundles intact. For example, in  Fig.~\ref{fig:len_dist_crop16}, where $h$ was set to ten while the ground truth was five, we observe that the major clusters remain well segmented and small clusters can be removed if they have too few elements (clusters 2,3,4,5 and 7 in Fig.~\ref{fig:len_dist_crop16} all have less than 10 elements). This is an appealing trait of the proximity based clustering, thus providing good results even when exact $h$ might be unknown.
The following parameters were fixed for all the tests. We set the reliable Hessian threshold $R_{h}$ to be 0.3 for all our tests. A $R_{h}$ of 0.0 would mean all points have reliable local orientation which would cause spurious MetaTracts detection. On the other hand a very high $ R_{h}$ would lead to be a decline in number of MetaTracts produced. The $\alpha$ and $\beta$ in $R_{h}$ are as explained in Frangi et al. \cite{Frangi1998} and set to 0.5. The length and the width parameters for the cylinders for the MetaTracts decide how coarse the fiber cylinders are. Larger cylinders will handle noisy local orientation better as it inspects a higher number of candidate points to extend the fiber. We used 10.00 and 2.00 for length and breadth, respectively for all tests. $\alpha$ and $\beta$ in equation~\ref{eqn:algo_1} decide how quickly the value of the factor decays, we have used [7-10] and half the length of cylinder (5.0). Our number of fiber bundle directions are limited, thus even for small $\eta$, the distinction between the orientation clusters is preserved quite well. We compared  $\eta=3-7$ experimentally without any dramatic change in results. 
%\begin{figure}
%\centering
% \subfloat[$\eta=7$]{\includegraphics[width=0.3\linewidth]{imagesMetaTracts/Rplot-ndim3k2}}
%  \subfloat[$\eta=3$]{\includegraphics[width=0.3\linewidth]{imagesMetaTracts/Rplot-ndim7k2}}
%  \caption{Embedding from $\mathbb{R}^m\to \mathbb{R}^\eta$, here m is the number of MetaTracts, 580 for this example. Followed by clustering into K=2 clusters.}
%  \label{fig:eta}
%\end{figure}
%\begin{figure}[htp]
%\centering
% \includegraphics[width=\linewidth]{imagesMetaTracts/kul-results.eps}
%  \caption{Results from orientation based dimensionality reduction and clustering. A is the original data set, B shows the results of fiber tracking. C shows the results of orientation based clustering and D shows the results of the proximity based clustering. }\label{fig:kull-full}
%\end{figure}

>>>>

\begin{figure}[tb]
\centering
	%\includegraphics[width=0.45\textwidth]{imagesMT2014/crop-16/length_dist.png}
		\includegraphics[width=0.45\textwidth,  trim = 0mm 100mm 0mm 50mm, clip]{imagesMT2014/image_length}
	\caption{(A) MetaTracts colored according to mean orientations. (B) length distribution of individual MetaTracts for a particular bundle.(unit for length is grid cube edge length)}
	\label{fig:length_distribution}
\end{figure}

@@

The directed Hausdorff distance is very sensitive to outliers in the data.
However, because MetaTracts after orientation clustering are constructed using cylinders with similar orientations, they are not plagued by outliers.
To cluster based on MetaTract proximity, we used single linkage hierarchical clustering.
Hierarchical clustering has a single parameter $h$, the desired number of clusters.
In hierarchical clustering each objects starts in its own cluster and clusters are merged based on some criterion.
Clusters are merged until there are only $h$ clusters left.
Hierarchical clustering is intuitive since it is easy to trace how clusters are formed and merged.
Single linkage clustering finds pairs of objects $p \in P$ and $q \in Q$ where $P \neq Q$ which are closer than other such pairs, and merges the containing clusters $P$ and $Q$.
We found that single linkage hierarchical clustering had two major drawbacks.
First, the clustering would produce some small clusters of just a few MetaTracts.
These MetaTracts were anomalies caused by noise in the data and did not represent true fiber bundles.
Second, if two fiber bundles were parallel for some of their length and then separated, they would sometimes be clustered into the same fiber bundle.
A short MetaTract which was parallel to both and did not extend into the separation region could form a link between the two fiber bundles, causing them to be clustered into a single bundle.
To address the problem of small clusters, we applied hierarchical clustering and then identified small clusters with few MetaTracts.
We removed the MetaTracts that were in those clusters from the data set and reapplied hierarchical clustering.
To address the problem of short MetaTracts joining different fiber bundles, we applied hierarchical clustering and then removed the
shortest fibers (fibers less than 0.6 times the median length) in each bundle. We then reapplied hierarchical clustering.
We repeated both until a steady state of clusters is reached and no new small fibers can be removed. The results of clustering the orientation clusters Fig.~\ref{fig:orientation_clustering}(A,C) are shown in Fig.~\ref{fig:orientation_clustering}(D,E) respectively. Note how the purple cluster Fig.~\ref{fig:orientation_clustering}(E) bundle ``7'' is extracted well.
\begin{figure}
\centering
		%\includegraphics[width=0.4\textwidth]{imagesMT2014/garb1}~
		\includegraphics[width=0.5\textwidth, trim = 0mm 70mm 0mm 20mm, clip]{imagesMT2014/image7b.pdf}
	%\includegraphics[width=0.45\textwidth]{imagesMT2014/crop-16/bundles}~ 
	%\includegraphics[height=0.35\textwidth]{imagesMT2014/crop-16/vol_mesh.png} 
	\caption{(A) Volume rendering of data set Figure~\ref{fig:data-char}A.(B) shows a single slice of the data set. (C) shows two of the extracted meshes together. (D) shows the meshes separately.}
	\label{fig:crop-16-decomp}
\end{figure}

@@

\section{Parameter choices}
\label{sec:param_choices}
\begin{figure}
\centering
	\includegraphics[width=0.5\textwidth,  trim = 10mm 40mm 0mm 15mm, clip]{imagesMT2014/image_length_histogram.pdf}
	\caption{Number of tracts and median, minimum and maximum length of individual MetaTracts in the orientation cluster Fig.~\ref{fig:orientation_clustering}, clustered into 10 clusters (Fig.~\ref{fig:crop-16-decomp}A top). The unit for length is the grid cube length. Figure~\ref{fig:length_distribution}(B) shows the length of individual MetaTracts in cluster 6. Clusters labeled 2,3,4,5 and 7 all have cardinality less than 10. }
	\label{fig:len_dist_crop16}
\end{figure}
The critical parameters are $K$ for the K-means in orientation based clustering and $h$ for the hierarchical clustering. K denotes the number of major directions, which is either known a priori or can be easily estimated by looking at the weaving pattern. 
Our framework is robust to the choice of $h$ (Section~\ref{sec:results}). Large $h$ keeps the major bundles intact. For example, in  Fig.~\ref{fig:len_dist_crop16}, where $h$ was set to ten while the ground truth was five, we observe that the major clusters remain well segmented and small clusters can be removed if they have too few elements (clusters 2,3,4,5 and 7 in Fig.~\ref{fig:len_dist_crop16} all have less than 10 elements). This is an appealing trait of the proximity based clustering, thus providing good results even when exact $h$ might be unknown.
The following parameters were fixed for all the tests. We set the reliable Hessian threshold $R_{h}$ to be 0.3 for all our tests. A $R_{h}$ of 0.0 would mean all points have reliable local orientation which would cause spurious MetaTracts detection. On the other hand a very high $ R_{h}$ would lead to be a decline in number of MetaTracts produced. The $\alpha$ and $\beta$ in $R_{h}$ are as explained in Frangi et al. \cite{Frangi1998} and set to 0.5. The length and the width parameters for the cylinders for the MetaTracts decide how coarse the fiber cylinders are. Larger cylinders will handle noisy local orientation better as it inspects a higher number of candidate points to extend the fiber. We used 10.00 and 2.00 for length and breadth, respectively for all tests. $\alpha$ and $\beta$ in equation~\ref{eqn:algo_1} decide how quickly the value of the factor decays, we have used [7-10] and half the length of cylinder (5.0). Our number of fiber bundle directions are limited, thus even for small $\eta$, the distinction between the orientation clusters is preserved quite well. We compared  $\eta=3-7$ experimentally without any dramatic change in results. 
%\begin{figure}
%\centering
% \subfloat[$\eta=7$]{\includegraphics[width=0.3\linewidth]{imagesMetaTracts/Rplot-ndim3k2}}
%  \subfloat[$\eta=3$]{\includegraphics[width=0.3\linewidth]{imagesMetaTracts/Rplot-ndim7k2}}
%  \caption{Embedding from $\mathbb{R}^m\to \mathbb{R}^\eta$, here m is the number of MetaTracts, 580 for this example. Followed by clustering into K=2 clusters.}
%  \label{fig:eta}
%\end{figure}
%\begin{figure}[htp]
%\centering
% \includegraphics[width=\linewidth]{imagesMetaTracts/kul-results.eps}
%  \caption{Results from orientation based dimensionality reduction and clustering. A is the original data set, B shows the results of fiber tracking. C shows the results of orientation based clustering and D shows the results of the proximity based clustering. }\label{fig:kull-full}
%\end{figure}

### 2014-09-22 metaTracts-pvis2014.tex

%% Abstract section.
\abstract{
%Carbon fiber reinforced polymers (CFRP) typically consist of two main components: the matrix and the reinforcement. 
%Most important in order to achieve the desired characteristics of a CFRP is the reinforcement component. 
%In case of CFRPs the reinforcements typically consist of bundles of endless carbon fibers woven into sheets of carbon fiber %cloth, which are then layered and combined to a CFRP by the matrix component. 
%For quality control of CFRP materials and components industrial 3D X-ray computed tomography (XCT) is increasingly applied %as it allows detailed characterizations of the material in a non-destructive manner.

This work introduces MetaTracts, a novel method for extracting and visualizing individual fiber bundles as well as their weaving pattern in endless carbon fiber reinforced polymer (CFRP) materials, scanned by X-ray computed tomography (XCT).

>>>>

%% Abstract section.
\abstract{
This work introduces MetaTracts, a novel method for extracting and visualizing individual fiber bundles as well as their weaving pattern in endless carbon fiber reinforced polymer (CFRP) materials, scanned by X-ray computed tomography (XCT).
